Analysis of Red Wine Data by Ryan Galvin
========================================================

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
library(ggplot2)
library(grid)
library(gridExtra)

```

```{r echo=FALSE, Load_the_Data}
# Load the Data
wine <- read.csv('wineQualityReds.csv')
```

> This data set contains a variety of chemical properties of red wine,
with ratings of each type of wine from 3 wine experts.

# Univariate Plots Section

> Let's look at a summary of our data to get an idea of what we're working with.

```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots}
summary(wine)

qplot(x = fixed.acidity, data = wine)
qplot(x = volatile.acidity, data = wine)
qplot(x = citric.acid, data = wine)
qplot(x = residual.sugar, data = wine)
qplot(x = chlorides, data = wine)
qplot(x = free.sulfur.dioxide, data = wine)
qplot(x = total.sulfur.dioxide, data = wine)
qplot(x = density, data = wine)
qplot(x = pH, data = wine)
qplot(x = sulphates, data = wine)
qplot(x = alcohol, data = wine)
qplot(x = quality, data = wine)


```



# Univariate Analysis



### What is the structure of your dataset?
> The dataset contains 12 observations of various properties of red wines. We can see from the histograms that some properties, like pH and density, follow an almost normal distribution, while others, like alcohol and volatile acidity are left or right skewed. The citric acid variable appears to be the most "messy" in the sense it doesn't really follow any sort of distribution.

### What is/are the main feature(s) of interest in your dataset?
> The main features of interest in this dataset are the relationships between each chemical property and the quality of the wine.

### What other features in the dataset do you think will help support your investigation into your feature(s) of interest?
> I will look into relationships between the input variables, and see for example whether or not the normally distributed variables have stronger correlation to quality of red wine over the non-normal distributions.




# Bivariate Plots Section

> I would like to look at the relationship between total and free sulfur dioxide, fixed and volatile acidity, and all of these vs. quality to start.

```{r echo=FALSE, Bivariate_Plots}
ggplot(aes(x = fixed.acidity, y = volatile.acidity), data = wine) + 
  geom_point(alpha = 0.25) +
  ggtitle("Fixed vs Volatile Acidity")

ggplot(aes(x = free.sulfur.dioxide, y = total.sulfur.dioxide), data = wine) + 
  geom_point(alpha = 0.25) +
  ggtitle("Free vs Total Sulfur Dioxide")

ggplot(aes(x = free.sulfur.dioxide, y = quality), data = wine) + 
  geom_jitter(alpha = 0.25) +
  ggtitle("Free Sulfur Dioxide and Quality of Red Wines")

ggplot(aes(x = total.sulfur.dioxide, y = quality), data = wine) + 
  geom_jitter(alpha = 0.25) +
  ggtitle("Total Sulfur Dioxide and Quality of Red Wines")

ggplot(aes(x = fixed.acidity, y = quality), data = wine) + 
  geom_jitter(alpha = 0.25) +
  ggtitle("Fixed Acidity and Quality of Red Wines")

ggplot(aes(x = volatile.acidity, y = quality), data = wine) +
  geom_jitter(alpha = 0.25) +
  ggtitle("Volatile Acidity and Quality of Red Wines")
```

# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the investigation. How did the feature(s) of interest vary with other features in the dataset?

> When we look at fixed acidity vs. volatile acidity, we don't see much of a relationship at all. Most wines have a fixed acidity around 8, and a volatile acidity between 0.4 and 0.8.

>However, we can see as free sulfur dioxide increases, total sulfur dioxide increases, which makes sense.

>Out of all of the quality plots, it looks like the features most likely to have a high quality wine rating are free sulfur dioxide and fixed acidity.


### Did you observe any interesting relationships between the other features (not the main feature(s) of interest)?

>The most interesting feature observed was that fixed acidity and volatile acidity have almost no correlation together.

### What was the strongest relationship you found?
>The strongest relationship I found was free sulfur dioxide to total sulfur dioxide, which has a pretty obvious linear monotonic relationship.

# Multivariate Plots Section

```{r echo=FALSE, Multivariate_Plots}
ggplot(aes(x = free.sulfur.dioxide, y = total.sulfur.dioxide, color = quality), data = wine) +
  geom_point(alpha = 0.50) +
  ggtitle("Free and Total Sulfur dioxide colored by Quality of Wine")

ggplot(aes(x = fixed.acidity, y = volatile.acidity, color = quality), data = wine) +
  geom_point(alpha = 0.50) +
  ggtitle("Fixed and Volatile acidity colored by Quality of Wine")

ggplot(aes(x = pH, y = density, color = quality), data = wine) + 
  geom_point(alpha = 0.50) +
  ggtitle("pH and Density colored by Quality of Wine")
  
ggplot(aes(x = sulphates, y = chlorides, color = quality), data = wine) +
  geom_point(alpha = 0.25) +
  ggtitle("Sulphates and Chlorides colored by Quality of Wine")
  
```

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the investigation. Were there features that strengthened each other in terms of looking at your feature(s) of interest?
> I looked at quality alongside my scatterplots of fixed and volatile acidity, and free and total sulfur dioxide. I found that higher quality wines tend to have lower free and total sulfur dioxide both, and also that higher quality wines tend to have lower volatile acidity, but don't really follow any sort of pattern for fixed acidity.

>I also decided to examine pH and Density as I thought it was interesting they were both roughly normally distributed. Higher quality wines tend to be a bit lower on the dense side, with most wines falling in the dead center of the pH and density scatterplot.

> I finally looked at chlorides and sulphates together colored by quality, and found that chlorides stay roughly the same for most wines, although with Sulphates, we see a variance in the quality of wine for slightly higher sulphates than lower sulphates.

### Were there any interesting or surprising interactions between features?

>I was surprised to see that fixed acidity could not be fit to any sort of pattern for quality, but volatile acidity could.

------

# Final Plots and Summary

> **Tip**: You've done a lot of exploration and have built up an understanding
of the structure of and relationships between the variables in your dataset.
Here, you will select three plots from all of your previous exploration to
present here as a summary of some of your most interesting findings. Make sure
that you have refined your selected plots for good titling, axis labels (with
units), and good aesthetic choices (e.g. color, transparency). After each plot,
make sure you justify why you chose each plot by describing what it shows.

### Plot One
```{r echo=FALSE, Plot_One}
qplot(x = quality, data = wine, binwidth = .25) +
  ggtitle("Histogram of Red Wine Qualities")
```

### Description One
>I choose to include the quality histogram so we can see where most wine quality falls, and thus be better able to interpret what would be considered a generally "high" quality of wine. We could say 6 and 7 are high enough quality wines for this data, given how few wines are rated an 8.

### Plot Two
```{r echo=FALSE, message=FALSE, warning=FALSE, Plot_Two}
ggplot(aes(x = free.sulfur.dioxide, y = total.sulfur.dioxide), data = wine) + 
  geom_point(alpha = 0.50) +
  ggtitle("Free vs Total Sulfur Dioxide of Red Wine") +
  xlab("Free Sulfur Dioxide (mg / dm^3)") +
  ylab("Total Sulfur Dioxide (mg / dm^3)") +
  geom_smooth()
```

### Description Two
>I include this scatterplot between free and total sulfur dioxide because I found the near linear relationship between them interesting, although expected.

### Plot Three
```{r echo=FALSE, Plot_Three}
ggplot(aes(x = fixed.acidity, y = volatile.acidity, color = quality), data = wine) +
  geom_point(alpha = 0.50) +
  xlab("Fixed Acidity (g / dm^3)") + 
  ylab("Volatile Acidity (g / dm^3") +
  ggtitle("Fixed and Volatile acidity colored by Quality of Wine")
```

### Description Three
> Finally, I include this quality colored scatterplot of the fixed and volatile acidity so we can see how the quality of the wine increases and decreases with our changes in the acidity levels. Higher quality wines tend to be associated with lower volatile acidity, whereas fixed acidity is a bit more spread out for the high quality wines.
------

# Reflection

> This red wine dataset has a lot of room for further exploration; I only initialled looked into sections that I found interesting. We could look at the other variables I didn't assess like sugar, citric acid, alcohol and more. 

